Generative AI for Image Reconstruction with Intensity Interferometry: A First Attempt

### 2024-10-02 Phase_II/draft/intro.tex
second edit

Intensity Interferometry (II) emerged in the 1970s as a groundbreaking technique for measuring the diameters, orbits, and limb-darkening coefficients of 32 stars in the N-body system. Despite these significant achievements, the method did not gain widespread adoption due to the then-required highly sensitive photon detectors and sophisticated data analysis techniques. However, advancements in computational methods and modern electronics, which now offer time resolution of detectors in the nanoseconds (ns) to picoseconds (ps) range, have renewed II as a feasible technique for resolving astrophysical objects at visible wavelengths. Current simulations with ns resolution of photon detectors show that II is effective in achieving high precision measurement of parameters for stellar objects \citep{10.1093/mnras/stab2391, 10.1093/mnras/stac2433}. However, a significant drawback of II is the loss of phase information. Consequently, a full reconstruction of the brightness distribution of a source with II necessitates phase information.

@@

Later \cite{holmes2010two} proposed an alternative method, utilizing the Cauchy-Riemann relations to reconstruct 1-D images and extending this to 2-D images across a range of signal-to-noise (SNR) values. This algorithm was applied to simulated data of stellar objects using II, considering existing and forthcoming Imaging Cherenkov Telescope Arrays (IACTs) with a large number of telescopes \citep{nunez2010stellar, nunez2012high, nunez2012imaging}.

In this paper, we propose an alternative approach using Generative Adversarial Networks (GANs) \citep{goodfellow2014generative} to reconstruct images of stellar objects using II. We consider four Imaging Cherenkov Telescope Arrays (IACTs) located in the Northern Hemisphere and perform a one-night simulation for a fast-rotating star. The predicted image using a trained GAN shows promising results for reconstructing the image's shape and size. The evaluation of the predicted image is conducted using image moments, demonstrating that the size and shape of the fast-rotating star are reconstructed. However, the brightness distribution recovers only at the third-order moment.

This paper is organized as follows: The first section discusses Intensity Interferometry following its signal and noise for fast-rotating stars along the Earth's rotation. The second section introduces the GAN formulation and its structure. Following the third section details the parameter selection for training the GAN to reconstruct the image. After that, the fourth section presents the results of trained GAN visually and using image moments. At the end, there is a discussion of the entire result.

>>>>

Intensity Interferometry (II) emerged in the 1970s as a groundbreaking technique for measuring the diameters, orbits, and limb-darkening coefficients of 32 stars in the N-body system. Despite these significant achievements, the method did not gain widespread adoption due to the then-required highly sensitive photon detectors and sophisticated data analysis techniques. However, advancements in computational methods and modern electronics, which now offer time resolution of detectors in the nanoseconds (ns) range, have renewed II as a feasible technique for resolving astrophysical objects at visible wavelengths. Current simulations with ns resolution of photon detectors show that II is effective in achieving high precision measurement of parameters for stellar objects \citep{10.1093/mnras/stab2391, 10.1093/mnras/stac2433}. Yet, despite these technological advancements, II still faces a significant limitation: the absence of phase information, which is essential for the complete reconstruction of the brightness distribution of stellar sources.

@@

Later \cite{holmes2010two} proposed an alternative method, utilizing the Cauchy-Riemann relations to reconstruct 1-D images and extending this to 2-D images across a range of signal-to-noise (SNR) values. This algorithm was applied to simulated data of stellar objects using II, considering existing and forthcoming Imaging Cherenkov Telescope Arrays (ICTAs) with a large number of telescopes \citep{nunez2010stellar, nunez2012high, nunez2012imaging}. Despite these theoretical advances, practical application to actual observational data of stellar objects with II has yet to be achieved. There is a need to advance this field by introducing a novel image reconstruction technique.

Currently, artificial intelligence (AI) is applied to high-volume data to reduce human computing resources and time effort. New analysis techniques using Machine Learning (ML) applications on observed data are playing at the front to reveal the astronomical mystery. In radio morphologies, the supervised and unsupervised algorithm is used to distinguish the radio components from the infrared source \citep{wu2019radio, galvin2020cataloguing}. Recently, the image reconstruction of the M87 black hole using Generative Adversarial Networks (GANs) has been performed \citep{10.1093/mnras/stad3797}. There is also a need to introduce the application of ML to II data, which will aim to overcome the limitations of image reconstruction and high-resolution optical astronomy.

In this paper, we propose an alternative approach for image reconstruction of stellar objects with II using GAN \citep{goodfellow2014generative}. We consider four ICTAs and perform a one-night simulation for a fast-rotating star. The predicted image by a trained GAN shows promising results for reconstructing the image's shape and size. Image moments evaluate the predicted image and demonstrate that the size and shape of the fast-rotating star are reconstructed. However, the brightness distribution recovers only up to the third-order moment. With the introduction of more telescopes, we can expect even greater precision, especially in recovering higher-order image moments. This approach holds great promise, pushing the boundaries of high-resolution optical astronomy and opening up new possibilities for studying stellar objects with remarkable detail.

This paper is organized as follows: The first section discusses Intensity Interferometry following its signal and noise for fast-rotating stars along the Earth's rotation. The second section introduces the GAN formulation and its structure. The third section details the parameter selection for training the GAN to reconstruct the image. After that, the fourth section presents the results of trained GAN visually and using image moments. At the end, there is a conclusion of the entire result.

### 2024-10-02 Phase_II/draft/main.tex
second edit

\title{Image reconstruction of stellar objects with Intensity Interferometry using Generative Adversarial Networks}
\date{}

@@

\begin{abstract}
In modern astronomy, the simulation of Intensity Interferometry demonstrates significant advancements in the high-precision resolution of stellar objects at optical wavelengths. Despite these achievements, phase retrieval remains challenging due to photon correlation, hence the image reconstruction. This paper proposes the application of a conditional Generative Adversarial Network (GAN) to Intensity Interferometry (II) data for high-resolution optical astronomy, aiming to overcome the limitations of traditional image reconstruction methods, as artificial intelligence (AI) techniques are at the forefront of revealing astronomical mysteries by analyzing observed data. Here, the cGAN effectively reconstructs the shape, size, and brightness distribution of a fast-rotating star by utilizing sparsely sampled, phase-subtracted Fourier transforms of the star. Simulations of II are based on an assembly of four telescopes similar to existing arrays. However, the sensitivity of the signal and high resolution are expected to improve with additional baselines. It makes the current and future Cherenkov Telescope Array Observatory (CTAO) an ideal candidate for II applications. This approach is highly relevant and innovative, addressing key challenges in phase reconstruction and proposing novel solutions that could revolutionize high-resolution imaging in astronomy.	
\end{abstract}
\input{intro}
\input{II}
\input{GAN}
\input{ParaGan}
\input{Result}
\input{conclusion}

>>>>

\title{Generative AI for Image Reconstruction with Intensity Interferometry: A First Attempt}
\date{}

@@

\begin{abstract}
In modern astronomy, the simulation of Intensity Interferometry demonstrates significant advancements in the high-precision resolution of stellar objects at optical wavelengths. Despite these achievements, phase retrieval of the signal remains challenging due to photon correlation, hence the image reconstruction. Nowadays, artificial intelligence (AI) techniques are at the forefront of revealing astronomical mysteries by analyzing observed data. This paper proposes the application of a conditional Generative Adversarial Network (cGAN) to Intensity Interferometry data for high-resolution optical astronomy, aiming to overcome the limitations of traditional image reconstruction methods. Here, the cGAN effectively reconstructs the shape, size, and brightness distribution of a fast-rotating star by utilizing sparsely sampled, phase-subtracted Fourier transforms of the star. Simulations of II are based on an assembly of four telescopes similar to existing arrays. However, the sensitivity of the signal and high resolution are expected to improve with additional baselines, which makes the current and future Cherenkov Telescope Array Observatory (CTAO) an ideal candidate for II applications. This approach is highly relevant and innovative, addressing key challenges in phase reconstruction and proposing novel solutions that could revolutionize high-resolution imaging in astronomy.	
\end{abstract}
\input{intro}
\input{II}
\input{GAN}
\input{ParaGan}
\input{Result}
\input{conclusion}
